---
title: "Resuscitation of microbial seed banks"
author: "Venus Kuo and Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

# Objective
Evaluate how recombinant Rpf affects plant traits

# Set working environment and load packages 

```{r setup, message = FALSE, warning = FALSE}
# Clear and set working directory 
rm(list = ls())
setwd("~/GitHub/BrassicaRpf/data")

# Require and/or install packages 
package.list <- c('vegan', 'plyr', 'grid', 'png', 'plyr', 'reshape2') 
for (package in package.list) {
  if (!require(package, character.only=T, quietly=T)) { 
    install.packages(package)
    library(package, character.only=T)
  } }

# Load sem function
sem <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}

# Load t-test custom functions # 
ttest <- function(reg, coefnum, val){
  co <- coef(summary(reg))
  tstat <- (co[coefnum,1]-val)/co[coefnum,2]
  pstat <- 2 * pt(abs(tstat), reg$df.residual, lower.tail = FALSE)
  return(list = c(t = tstat, df = reg$df.residual, p =  pstat))
}
```

# Rpf effects on microbial abundance

```{r}
```{r}
# Load microbial qPCR data
abundance <- read.csv("~/GitHub/BrassicaRpf/data/qPCR.txt", sep = ",", header = TRUE) 

# remove week 1 data 
abundance <- subset(abundance, Week == "6")

# Reset data frame index 
rownames(abundance) <- NULL

# Standarde gene copy abundances by soil 
abundance$stdGeneCopy <- (abundance$GeneCopy)/(abundance$Soil)

# Log10 transform standardized gene copy number 
abundance$log10stdGeneCopy <- log10(abundance$stdGeneCopy)

# Split abundance dataset by fungi and bacteria  
data.sub.bac <- subset(abundance, Gene == "rRNA")
data.sub.fun <- subset(abundance, Gene == "ITS")

# Calculate fungi to bacteria ratios 
gca.B <- data.sub.bac$stdGeneCopy
gca.F <- data.sub.fun$stdGeneCopy
abundance$FB <- (gca.F/gca.B)
ratioFB <- abundance[1:18, ]
ratioFB <- subset(ratioFB, select=(c(Treatment, Week, FB)))

# Subset abundance dataset 
data.sub <- subset(abundance, select = (c(Treatment, Gene, stdGeneCopy)))

# Melting the dataset 
data.m <- melt(data.sub)

# Bacterial abundance data for plotting 
bac.rpf <- data.m[ which(data.m$Treatment == "Rpf+" & data.m$Gene == "rRNA"),]
bac.con <- data.m[ which(data.m$Treatment == "Rpf-" & data.m$Gene == "rRNA"),]

# Bacterial abundance summary table 
bac.mean <- aggregate(data.sub.bac$stdGeneCopy ~ Treatment, data.sub.bac, mean)
bac.sem <- aggregate(data.sub.bac$stdGeneCopy ~ Treatment, data.sub.bac, sem)
bac.sem.LL <- bac.mean[2] - bac.sem[2]
bac.sem.UL <- bac.mean[2] + bac.sem[2]
bac.table <- data.frame(bac.mean[1], bac.mean[2], bac.sem[2],
          bac.sem.LL[1], bac.sem.UL[1])
colnames(bac.table) <- c("Treatment", "mean", "sem", "LCI", "UCI")
bac.table <- bac.table[order(bac.table$mean),]

# Fungal abundance data for plotting 
fun.rpf <- data.m[ which(data.m$Treatment == "Rpf+" & data.m$Gene == "ITS"),]
fun.con <- data.m[ which(data.m$Treatment == "Rpf-" & data.m$Gene == "ITS"),]

# Fungal abundance summary table
fun.mean <- aggregate(data.sub.fun$stdGeneCopy ~ Treatment, data.sub.fun, mean)
fun.sem <- aggregate(data.sub.fun$stdGeneCopy ~ Treatment, data.sub.fun, sem)
fun.sem.LL <- fun.mean[2] - fun.sem[2]
fun.sem.UL <- fun.mean[2] + fun.sem[2]
fun.table <- data.frame(fun.mean[1], fun.mean[2], fun.sem[2],
          fun.sem.LL[1], fun.sem.UL[1])
colnames(fun.table) <- c("Treatment", "mean", "sem", "LCI", "UCI")
fun.table <- fun.table[order(fun.table$mean),]

# Fungi: Bacteria ratio data for plotting
FB.rpf <- ratioFB[ which(ratioFB$Treatment == "Rpf+"),]
FB.con <- ratioFB[ which(ratioFB$Treatment == "Rpf-"),]

# Fungal: Bacteria ratio summary table
FB.mean <- aggregate(ratioFB$FB ~ Treatment, ratioFB, mean)
FB.sem <- aggregate(ratioFB$FB ~ Treatment, ratioFB, sem)
FB.sem.LL <- FB.mean[2] + FB.sem[2]
FB.sem.UL <- FB.mean[2] - FB.sem[2]
FB.table <- data.frame(FB.mean[1], FB.mean[2], FB.sem[2],
          FB.sem.LL[1], FB.sem.UL[1])
colnames(FB.table) <- c("Treatment", "mean", "sem", "LCI", "UCI")
FB.table <- FB.table[order(FB.table$mean),]
```

# Statistical tests

```{r}
# Bacterial abundance
bac.ttest <- t.test(stdGeneCopy ~ Treatment, data = data.sub.bac)
bac.ttest    # Significant: t = 2.7069, df = 14.777, p = 0.0164

# Fungal abundance 
fun.ttest <- t.test(stdGeneCopy ~ Treatment, data = data.sub.fun)
fun.ttest    # Significant: t = -3.2719, df = 11.701, p = 0.00689

# F:B ratio
FB.ttest <- t.test(FB ~ Treatment, data = ratioFB)
FB.ttest     # Significant: t= -2.8454, df = 9.4476, p = 0.01835
```


# Plot microbial abundance

```{r}
png(filename="../figures/Fig4.MicroAbund.png",
    width = 1200, height = 800, res = 96*2)
par(oma=c(7,3,7,1), mar=c(2,3,3,4), mfrow=c(1,3))

# Panel 1: Bacterial abundance 
abun.bac.fig <- plot(jitter(rep(1, length(bac.con$value)), amount = 0.1), bac.con$value, 
      ylim = c(0, 2.4E6), xlim = c(0.5, 2.5), pch = 21, col = "lightgrey", bg = "lightgrey", lwd = 3.5,
      cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, cex.axis = 2,
      las = 1, ylab = "", xlab = "")
      box(lwd = 2)
points(jitter(rep(2, length(bac.rpf$value)), amount = 0.1), bac.rpf$value, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)

# Add means
points(1, mean(bac.con$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(2, mean(bac.rpf$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  

box(lwd = 2)

# Y axis labels
mtext(expression('16S rRNA gene copy/ g soil'), side = 2,
      outer = FALSE, cex = 1, line = 3.5, adj = 0.5)

# Major axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1, 
     labels = c("0.0E0", "8.0E5", "1.6E6", "2.4E6"), 
     at = c(0.0E0, 8.0E5, 1.6E6, 2.4E6))

axis(side = 4, lwd.ticks = 2, cex.axis = 1, las = 1,
     at=c(0.0E0, 8.0E5, 1.6E6, 2.4E6), labels = F, tck = -0.05)

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("-Rpf", "+Rpf"), at = c(1, 2))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     at = c(1, 2), labels = F, tck = -0.05)

# Add SEM
arrows(x0 = c(2,1), y0 = bac.table$mean, y1 = bac.table$LCI, angle = 90,
       length = 0.15, lwd = 2)
arrows(x0 = c(2,1), y0 = bac.table$mean, y1 = bac.table$UCI, angle = 90,
       length=0.15, lwd = 2)

# Panel label
text(0.7,2.3E6, labels = "A", col="black", cex = 2)

# P-value label
mtext(text = expression(italic("P")~" = 0.016") , side = 3, line = -1.2, 
      adj = 0.925, col = "black", cex = 0.8)
# Sample number label
#mtext(text = expression(italic("n")~" = 9"), side = 3, line = -2.2, adj = 0.925, col="black", cex=0.8)

# Panel 2: fungal abundance
abun.fun.fig <- plot(jitter(rep(1, length(fun.con$value)), amount = 0.1), 
      fun.con$value, ylim = c(0, 1.5E6), xlim = c(0.5, 2.5), pch = 21, 
      col = "lightgrey", bg = "lightgrey", lwd = 3.5, cex = 1.7, yaxt = "n", 
      xaxt = "n", cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")
      box(lwd = 2)
points(jitter(rep(2, length(fun.rpf$value)), amount = 0.1), fun.rpf$value, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)

# Add means 
points(1, mean(fun.con$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(2, mean(fun.rpf$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  

box(lwd = 2)

# Y axis labels
mtext(expression('ITS gene copy / g soil'), side = 2,
      outer = FALSE, cex = 1, line = 3.5, adj = 0.5)

# Major axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1, 
     labels = c("0.0E0", "5.0E5", "1.0E6", "1.5E6"), 
     at = c(0.0E0, 5.0E5, 1.0E6, 1.5E6))

axis(side = 4, lwd.ticks = 2, cex.axis = 1, las = 1,
     at=c(0.0E0, 5.0E5, 1.0E6, 1.5E6), labels = F, tck = -0.05)

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("-Rpf", "+Rpf"), at = c(1, 2))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     at = c(1, 2), labels = F, tck = -0.05)

# Add SEM
arrows(x0 = c(1,2), y0 = fun.table$mean, y1 = fun.table$LCI, angle = 90,
       length = 0.15, lwd = 2)
arrows(x0 = c(1,2), y0 = fun.table$mean, y1 = fun.table$UCI, angle = 90,
       length=0.15, lwd = 2)

# Panel label 
text(0.7, 1.45E6, labels = "B", col = "black", cex = 2)

# p-value 
mtext(text = expression(italic("P")~" = 0.007") , side = 3, line = -1.2, 
      adj = 0.925, col = "black", cex = 0.8)
# Sample number label
#mtext(text = expression(italic("n")~" = 9"), side = 3, line = -2.2, adj = 0.925, col="black", cex=0.8)

# Panel 3: F:B
FB.fig <- plot(jitter(rep(1, length(FB.con$FB)), amount = 0.1), FB.con$FB, 
      ylim = c(0, 1.5), xlim = c(0.5, 2.5), pch = 21, col = "lightgrey", 
      bg = "lightgrey", lwd = 3.5, cex = 1.7, yaxt = "n", xaxt = "n", 
      cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")
      box(lwd = 2)
points(jitter(rep(2, length(FB.rpf$FB)), amount = 0.1), FB.rpf$FB, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)

# Adding mean data pointfor each treatment # 
points(1, mean(FB.con$FB), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(2, mean(FB.rpf$FB), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  

box(lwd = 2)

# Y axis labels
mtext(expression('Fungal : bacterial ratio'), side = 2,
      outer = FALSE, cex = 1, line = 2.5, adj = 0.5)

# Major Axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1, 
     labels = c("0.0", "0.5", "1.0", "1.5"), 
     at = c(0.0, 0.5, 1.0, 1.5))

axis(side = 4, lwd.ticks = 2, cex.axis = 1, las = 1,
     at=c(0.0, 0.5, 1.0, 1.5), labels = F, tck = -0.05)

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("-Rpf", "+Rpf"), at = c(1, 2))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     at = c(1, 2), labels = F, tck = -0.05)

# Add SEM 
arrows(x0 = c(1,2), y0 = FB.table$mean, y1 = FB.table$LCI, angle = 90,
       length = 0.15, lwd = 2)
arrows(x0 = c(1,2), y0 = FB.table$mean, y1 = FB.table$UCI, angle = 90,
       length = 0.15, lwd = 2)

# Panel label
text(0.7, 1.45, labels = "C", col = "black", cex = 2)

# p-value 
mtext(text = expression(italic("P")~" = 0.018") , side =3, line = -1.2, 
      adj = 0.925, col = "black", cex = 0.8)
# Sample number label
#mtext(text = expression(italic("n")~" = 9"), side = 3, line = -2.2, adj = 0.925, col="black", cex=0.8)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("../figures/Fig4.MicroAbund.png")
grid.raster(img)
```