---
title: "DormancyGPCA - AbundanceDiversity"
author: "Venus Kuo"
date: "November 13th, 2017"
output: pdf_document
---

# 1) Questions

1. We test whether a concentration gradent of recombiant KBS0714 Rpf protein can alter the diversity and abundance of cultured soil bacteria on NA Petri plates.


# 2) Set working environment and load packages

```{r setup}
# Set working directory #
rm(list = ls())
getwd()
knitr::opts_knit$set(root.dir = "~/../Github/BrassicaRpf/data")

# Require or install packages #
package.list <- c('vegan', 'ggplot2', 'gplots', 'plyr', 'psych', 'corrplot', 'BiodiversityR', 'RColorBrewer') 
for (package in package.list) {
  if (!require(package, character.only=T, quietly=T)) { 
    install.packages(package)
    library(package, character.only=T)
} }
```


# 3) Describe, load, and modify data set

plateCfuMorphotypeData.txt : A text file containing the soil sample location, treatment concentration, replicate number, dilution factor, colony counts, and morphology counts. 

Data set modification: 
1. I calulated colony forming units (CFU) based on colony counts and dilution factor. 
2. I standardized CFU and colony morphotypes  by soil amount (g). 
3. I log10 transformed standardized CFU and morphotype.
4. I log2 transformed the concentration gradient of Rpf treatment due to uneven concentration treatments. 

```{r}
# Load plateCfuMorphotypeData.txt dataset #
cfu <- read.csv("plateCfuMorphotypeData.txt",  header=T)

# Calculating and standardizing CFU by soil amount # 
colony <- cfu$Colonies
dil.factor <- cfu$Dilution
soilamount <- cfu$Soil
cfu$CFU <- colony/dil.factor
cfu$stdCFU <- cfu$CFU/soilamount
cfu$stdcfu.log<- log10(cfu$stdCFU) 

# Standardizing unique colony morphotype # 
uniq <- cfu$Morphotypes
cfu$stdMorph <- uniq/soilamount
cfu$stdMorph.log <- log10(cfu$stdMorph)

# Natural log transform Rpf concentration # 
rpf.conc <- cfu$RpfConcentration
cfu$rpf <- log2(rpf.conc + 1) # I added +1 to make natural log of 0 as 1 

#View(cfu) # Look at the data set 
```


# 4) Scatterplot visualiation of cultured CFU and morphotypes along Rpf gradient

## Color coding each soil sample by color

```{r}
# Create a custom color scale for each soil sample #
myColors <- brewer.pal(3,"Set1")  

# Assign the three soil samples (Peru, Dunn, Indiana Preserves) to each color # 
names(myColors) <- levels(cfu$Location)

# Assign soil sample color to colScale # 
colScale <- scale_colour_manual(name = "Location", values = myColors)
```


## Line graph visualization of log transformed cultured soil bacterial abundance standardized by soil

```{r}
# Calculate the average and standard error of means of log transformed standardized CFU  # 
cfu.mean <- ddply(cfu, c("rpf", "Location"), summarise, 
                   mean=mean(stdcfu.log), sem=sd(stdcfu.log)/sqrt(length(stdcfu.log)))
cfu.mean <- transform(cfu.mean, lower=mean-sem, upper=mean+sem) 

# Line graph # 
cfu.line <- ggplot(data=cfu.mean, aes(x=rpf, y=mean, colour=Location)) + 
  geom_line()

# Edit line graph figure # 
cfu.plot <- cfu.line + scale_y_continuous(limits = c(8.0, 9.5), breaks = seq(8.0, 9.5, 0.5), 
                     sec.axis = sec_axis(~ . * 1, labels = c(" "," "," ", " "))) +
    geom_line(size=1.25) +
  labs(x="Rpf concentration,"~Log["2"](paste(mu)~mol/L),
       y=Log["10"]~(CFU~g^-1~ soil)) +
  geom_errorbar(aes(ymax=upper, ymin=lower), data=cfu.mean, width = 0.2, size = 0.5) +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=20, margin =margin(0,15,0,0)),
        axis.text.x=element_text(colour="black", size =20, margin =margin(15,0,0,0)),
        axis.title.y = element_text(colour="black", size=20,  margin = margin(0,20,0,10)),
        axis.title.x = element_text(colour="black", size=20,  margin=margin(20,0,0,0)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
          axis.ticks = element_line(size = 1.5),
          axis.ticks.length = unit(.25, "cm"),
        legend.title = element_text(size=15),
        legend.text=element_text(size=15)) +
  scale_x_continuous(breaks = round(seq(0, 6, by = 1),1)) + 
  geom_point(size = 3)

# Plot CFU line graph # 
cfu.plot
```


## Scatter plot visualization of log transformed cultured soil diversity standardized by soil amount

```{r}
# Calculate the average and standard error of means of log transformed standardized morphotype  # 
uniq.mean <- ddply(cfu, c("rpf", "Sample"), summarise, 
                   mean=mean(stdMorph), sem=sd(stdMorph)/sqrt(length(stdMorph)))
uniq.mean <- transform(uniq.mean, lower=mean-sem, upper=mean+sem) 

# Line graph morphotype # 
uniq.line <- ggplot(data=uniq.mean, aes(x=rpf, y=mean, colour=Sample)) + 
  geom_line()

# Edit line graph figure # 
uniq.plot <- uniq.line +  
  geom_line(size=1.25) +
   labs(x=expression(paste("Rpf concentrations Log2(",mu,"mol/L)")), 
       y="Morphotype/Soil (g)") +
  geom_errorbar(aes(ymax=upper, ymin=lower), data=uniq.mean, width = 0.2, size = 0.5) +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=20, margin =margin(0,15,0,0)),
        axis.text.x=element_text(colour="black", size =20, margin =margin(15,0,0,0)),
        axis.title.y = element_text(colour="black", size=20,  margin = margin(0,20,0,10)),
        axis.title.x = element_text(colour="black", size=20, margin=margin(20,0,0,0)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(-0.35, "cm"),
        axis.ticks = element_line(size = 1.65, colour="black"),
        legend.title = element_text(size=15),
        legend.text=element_text(size=15)) +
  scale_x_continuous(breaks = round(seq(0, 6, by = 1),1)) +
  geom_point(size = 3) +
  annotate("text", x =0.1, y=37, label = c("B"), size=12)

# Plot Line graph of morphotype # 
uniq.plot  


```


# 5) Fitting quadratic curve to data set

## Soil bacterial CFU

```{r}
linear.model <-lm(stdcfu.log ~ rpf, cfu)
summary(linear.model)

file <- paste("~/../GitHub/DormancyGPCA/figure/CFU_vs_Rpf.png", sep="")
png(filename=file, width=5, height=5, units="in", res=600, pointsize=10)

plot.new()
par(mfrow=c(1, 1), mar = c(5,5,3,1), oma =c(0,0,2,0))

mod <- lm(stdcfu.log ~ rpf + I(rpf^2), data=cfu)
dat2 <- data.frame(cbind('rpf'=cfu$rpf, 'cfu'=predict(mod)))
dat2 <- dat2[order(dat2), ]
dat2 <- dat2[complete.cases(dat2), ]

r2 <- summary(mod)$r.squared
coefs <- summary(mod)$coefficients
summary(mod)

title <- paste("Colonies vs Rpf,\n r-square=", round(r2,2))
plot(cfu$rpf, cfu$stdcfu.log, main=title, 
  	xlab="Rpf concentration, log2(umol/L)", ylab="Log10(CFU/Soil(g))",
  	cex.lab=1.5, cex.main=1.5, pch=19)


# predicts + interval
newx <- seq(min(cfu$rpf), max(cfu$rpf), len=100)
preds <- predict(mod, data.frame(rpf = newx), interval = "confidence") 

clr <- 'gray' #c(2,2,3,4,5) 
clr <- adjustcolor(clr, alpha.f = 0.3) 
polygon(c(rev(newx), newx), c(rev(preds[ ,3]), preds[ ,2]), col = clr, border = NA)

lines(dat2$rpf, dat2$cfu, col="red", lwd=2) # regression line (y~x) 
lines(lowess(dat2$rpf, dat2$cfu), col="blue", lwd=2) # lowess line (x,y)
points(cfu$rpf, cfu$cfu, pch=19)
dev.off()



```


## Plot morphotypes 
```{r}
# Perform linear model # 
linear.model.u <- lm(stdMorph ~rpf, cfu)
summary(linear.model.u)

# Make quadratic curve fit figure # 
file <- paste("~/../GitHub/DormancyGPCA/figure/Morphotypes_vs_Rpf.png", sep="")
png(filename=file, width=5, height=5, units="in", res=600, pointsize=10)

plot.new()
par(mfrow=c(1, 1), mar = c(5,5,3,1), oma =c(0,0,2,0))

# Quadratic curve #
mod <- lm(stdMorph ~ rpf + I(rpf^2), data=cfu)
#mod <- lm(stdMorph ~ rpf, cfu)
dat2 <- data.frame(cbind('rpf'=cfu$rpf, 'stdMorph'=predict(mod)))
dat2 <- dat2[order(dat2), ]
dat2 <- dat2[complete.cases(dat2), ]

r2 <- summary(mod)$r.squared
coefs <- summary(mod)$coefficients
summary(mod)

title <- paste("Morphotypes vs Rpf,\n r-square=", round(r2,2))
plot(cfu$rpf, cfu$stdMorph, main=title, 
  	xlab="Rpf concentration, log2(umol/L)", ylab="Morphotypes/Soil (g)",
  	cex.lab=1.5, cex.main=1.5, pch=19)

# predicts + interval
newx <- seq(min(cfu$rpf), max(cfu$rpf), len=100)
preds <- predict(mod, data.frame(rpf = newx), interval = "confidence") 

clr <- 'gray' #c(2,2,3,4,5) 
clr <- adjustcolor(clr, alpha.f = 0.3) 
polygon(c(rev(newx), newx), c(rev(preds[ ,3]), preds[ ,2]), col = clr, border = NA)

lines(dat2$rpf, dat2$stdMorph, col="red", lwd=2) # regression line (y~x) 
lines(lowess(cfu$rpf, cfu$stdMorph), col="blue", lwd=2) # lowess line (x,y)
points(cfu$rpf, cfu$stdMorph, pch=19)
dev.off()

```


# Generalized Linear Models # 

```{r}
# Poisson Regression # 
# Used when predicting an outcome variable representing counts from a set of continuous predictor variables #
fit.cfu <- glm(stdcfu.log ~ rpf , data= cfu , family=quasipoisson())
summary(fit.cfu)

fit.uc <- glm(stdMorph~rpf,data=cfu,family=poisson())
summary(fit.uc)
```







