---
title: "Resuscitation of microbial seed banks"
author: "Venus Kuo and Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

# 1. Objective
Evaluate how recombinant Rpf affects Actinobacteria diversity and composition

## Set working environment and load packages 

```{r setup, message = FALSE, warning = FALSE}
# Clear and set working directory 
rm(list = ls())
setwd("~/GitHub/BrassicaRpf/data")

# Require and/or install packages 
package.list <- c('vegan', 'nlme' ,'data.table', 'plyr', 'pander', 'reshape', 
                  'grid', 'png', 'car', 'bbmle', 'reshape2', 'ggplot2', 
                  'picante', 'ade4', 'phangorn', 'indicspecies', 'reshape') 
for (package in package.list) {
  if (!require(package, character.only=T, quietly=T)) { 
    install.packages(package)
    library(package, character.only=T)
  } }

# Load sem function
sem <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
```


# 2. Load all data 

```{r, message=FALSE, warning=FALSE}
source("~/GitHub/BrassicaRpf/bin/DiversityFunctions.R")
source("~/GitHub/BrassicaRpf/bin/MothurTools2.R")
source("~/GitHub/BrassicaRpf/bin/phylodiversity2.R")

# Run All: Select if all section are to be re-run
run.all <- TRUE

# Load files
  # design = general design file for experiment
  # shared = OTU table from mothur with sequence similarity clustering
  # tax    = Taxonomy for 97% similarity OTUs

design <- "~/GitHub/BrassicaRpf/data/Brassica.design.txt"
shared <- "~/GitHub/BrassicaRpf/mothur/output/Brassica.bac.final.shared"
tax <- "~/GitHub/BrassicaRpf/mothur/output/Brassica.bac.final.0.03.taxonomy"

# Import Design 
design <- read.delim(design, header=T, row.names=1)
# Import Shared Files
OTU <- read.otu(shared = shared, cutoff = "0.03") # 97% similarity 
# Import Taxonomy
OTU.tax <- read.tax(taxonomy = tax, format = "rdp")  
# Import Phylogenetic tree 
OTU.tre <- read.tree("../phylo/Brassica.bac.rename.tree.2") 
```

# 3. Subset Actinobacteria

```{r}
# Actinobacteria taxonomy file
Actino.OTU.tax <- OTU.tax[ which(OTU.tax$Phylum == 'Actinobacteria'), ] # = 2724 OTU
Actino.OTU <- OTU[, Actino.OTU.tax$OTU]
# Remove OTUs with less than two occurrences across all sites # 
Actino.OTU <- Actino.OTU[, which(colSums(Actino.OTU) >= 2)]
# Remove mock community # 
Actino.OTU <- Actino.OTU[1:20, ]
```

# 4. Alpha diversity

```{r}
# Resampling code to estimate alpha diversity (used if run.all = T)
if (run.all == TRUE){
  rich <- round(richness.iter(input = Actino.OTU, size = 950,
                              iters = 100, shared = "FALSE"), 3)
  even <- round(evenness.iter(input = Actino.OTU, size = 950,
                              iters = 100, shared = "FALSE",
                              method = "Evar"), 3)
  rare <- rarefy(Actino.OTU, 950, se = FALSE, MARGIN = 1)
  # Write output to files
  write.table(rich, "../data/rich.txt", sep = "\t",
              col.names = T, row.names = T)
  write.table(even, "../data/even.txt", sep = "\t",
              col.names = T, row.names = T)
}

# Read in alpha diversity files from above
rich2 <- read.table("../data/rich.txt", sep = "\t")
even2 <- read.table("../data/even.txt", sep = "\t")

# Merge data to design and calculate mean and sem per sample
rich.data <- merge(design, rich2, by = "row.names")
row.names(rich.data) <- rich.data$Row.names
rich.data <- rich.data[sort(row.names(rich.data)), ]
rich.mean <- round(apply(rich.data[5:(4 + dim(rich2)[2])], 1, mean, na.rm = TRUE),3)
rich.sem <- round(apply(rich.data[5:(4 + dim(rich2)[2])], 1, sem, na.rm = TRUE), 3)

even.data <- merge(design, even2, by = "row.names")
row.names(even.data) <- even.data$Row.names
even.data <- even.data[sort(row.names(even.data)), ]
even.mean <- round(apply(even.data[5:(4 + dim(even2)[2])], 1, mean, na.rm = TRUE),3)
even.sem <- round(apply(even.data[5:(4 + dim(even2)[2])], 1, sem, na.rm = TRUE),4)

# Make new dataframe merging design file and mean diversity
Brassica.div <- data.frame(design[sort(row.names(design)), ], rich.mean, even.mean)

# Take averages of technical reps
rich.rep.ave <- ddply(Brassica.div, .(treatment, type, rep), summarize, rich = mean(rich.mean))
even.rep.ave <- ddply(Brassica.div, .(treatment, type, rep), summarize, even = mean(even.mean))

# Reshape data 
rich.2 <- reshape(rich.rep.ave[,1:4], timevar = "type",
                   idvar = c("treatment", "rep"), direction = "wide")

even.2 <- reshape(even.rep.ave[,1:4], timevar = "type",
                   idvar = c("treatment", "rep"), direction = "wide")

rich.means <- data.frame(mean(rich.2[1:5,3]), mean(rich.2[6:10,3]), 
                         mean(rich.2[1:5,4]), mean(rich.2[6:10,4]))
colnames(rich.means) <- c("-Rpf,RNA", "+Rpf,RNA", "-Rpf,DNA", "+Rpf,DNA")

rich.sem <- data.frame(sem(rich.2[1:5,3]), sem(rich.2[6:10,3]), 
                         sem(rich.2[1:5,4]), sem(rich.2[6:10,4]))
colnames(rich.sem) <- c("-Rpf,RNA", "+Rpf,RNA", "-Rpf,DNA", "+Rpf,DNA")


even.means <- data.frame(mean(even.2[1:5,3]), mean(even.2[6:10,3]), 
                         mean(even.2[1:5,4]), mean(even.2[6:10,4]))
colnames(even.means) <- c("-Rpf,RNA", "+Rpf,RNA", "-Rpf,DNA", "+Rpf,DNA")

even.sem <- data.frame(sem(even.2[1:5,3]), sem(even.2[6:10,3]), 
                         sem(even.2[1:5,4]), sem(even.2[6:10,4]))
colnames(even.sem) <- c("-Rpf,RNA", "+Rpf,RNA", "-Rpf,DNA", "+Rpf,DNA")

# Statistical test

# One-way ANOVA
RNA <- Brassica.div[ which(Brassica.div$type == 'cDNA'), ]
DNA <- Brassica.div[ which(Brassica.div$type == 'DNA'), ]

RNA.aov.rich <- aov(rich.mean ~ treatment, RNA)
summary(RNA.aov.rich)

DNA.aov.rich <- aov(rich.mean ~ treatment, DNA)
summary(DNA.aov.rich)

RNA.aov.even <- aov(even.mean ~ treatment, RNA)
summary(RNA.aov.even)

DNA.aov.even <- aov(even.mean ~ treatment, DNA)
summary(DNA.aov.even)

# Two-way ANOVA
aov2way.rich <- aov(rich.mean ~ treatment*type, Brassica.div)
summary(aov2way.rich)

aov2way.even <- aov(even.mean ~ treatment*type, Brassica.div)
summary(aov2way.even)
```

## 5. Beta diversity

```{r}
# Make presence-absence matrix
Actino.OTU.PA <- (Actino.OTU > 0) * 1

# Make relative abundance matrix
Actino.OTU.REL <- Actino.OTU
for (i in 1:dim(Actino.OTU)[1]){
  Actino.OTU.REL[i,] <- Actino.OTU[i,]/sum(Actino.OTU[i,])
  }

# Log-transform relative abundances
Actino.OTU.REL.log <- decostand(Actino.OTU, method="log")

Brassica.bc.dis <- vegdist(Actino.OTU.REL.log, method = "bray", binary = "FALSE")
Brassica.dis.mean <- mean(Brassica.bc.dis)

# Principal Coordinates Analysis (PCoA)
Brassica.PCoA <- cmdscale(Brassica.bc.dis, eig = TRUE, k = 3)
explainvar1 <- round(Brassica.PCoA$eig[1] / sum(Brassica.PCoA$eig), 3) * 100 # 23.7
explainvar2 <- round(Brassica.PCoA$eig[2] / sum(Brassica.PCoA$eig), 3) * 100 # 18.7
explainvar3 <- round(Brassica.PCoA$eig[3] / sum(Brassica.PCoA$eig), 3) * 100 # 8.8
sum.eig <- sum(explainvar1, explainvar2, explainvar3) # 51.2

# Actino.OTU Scores
Actino.OTU.scores <- t(cor(Brassica.PCoA$points, Actino.OTU.REL))
Actino.OTU.scores <- as.matrix(Actino.OTU.scores)[,1:2]
Actino.OTU.scores <- Actino.OTU.scores[abs(Actino.OTU.scores[,1]) > 0.7|abs(Actino.OTU.scores[,2]) > 0.7,]

# Average BC Distance Between Treatments
Brassica.bc.dis.m <- as.matrix(Brassica.bc.dis)
all.equal(row.names(Brassica.div), rownames(Brassica.bc.dis.m))

treatment.div <- unique(Brassica.div$treatment)
treatment.dis <- rep(NA, length(treatment.div))
for(i in 1:length(treatment.div)){
  temp <- row.names(Brassica.div[Brassica.div$treatment == treatment.div[i], ])
  treatment.dis[i] <- Brassica.bc.dis.m[temp[1], temp[2]]
}

#mean(treatment.dis)
```

# 6. Ordination

```{r}
# Plot figure -- Supplement for all bacteria ordination
png(filename="../figures/FigS5.ActinobacteriaComp.png",
    width = 1800, height = 800, res = 96*2)

layout(matrix(1:3, 1, 3), widths = c(20, 9, 2.5))

par(mar = c(7, 10, 1, 0) + 0.5)

plot(Brassica.PCoA$points[ ,1], Brassica.PCoA$points[ ,2],
     ylim = c(-0.475, 0.45), xlim = c(-0.5, 0.4),
     xlab = paste("PCoA 1 (", explainvar1, "%)", sep = ""),
     ylab = paste("PCoA 2 (", explainvar2, "%)", sep = ""), line = 5,
     #xlab = "", ylab = "", xaxt = "n", yaxt = "n",
     pch = 22, cex = 2.0, type = "n", cex.lab = 2.5, cex.axis = 3,
     axes = FALSE)

# Add Axes
axis(side = 1, labels = T, lwd.ticks = 2, cex.axis = 2, las = 1, tck=-0.025)
axis(side = 2, labels = T, lwd.ticks = 2, cex.axis = 2, las = 1, tck=-0.025)
axis(side = 3, labels = F, lwd.ticks = 2, cex.axis = 2, las = 1, tck=-0.025)
axis(side = 4, labels = F, lwd.ticks = 2, cex.axis = 2, las = 1, tck=-0.025)
abline(h = 0, v = 0, lty = 3)

box(lwd = 2)

# Subset data
all.equal(row.names(Brassica.PCoA$points), rownames(Brassica.div))
Brassica.points <- data.frame(Brassica.PCoA$points, Brassica.div)

# Active community
Brassica.active.rpf <- Brassica.points[ which(Brassica.points$type == "cDNA" &
                                   Brassica.points$treatment == "Rpf+"), ]
Brassica.active.no <- Brassica.points[ which(Brassica.points$type == "cDNA" &
                                   Brassica.points$treatment == "Rpf-"), ]
# Total community
Brassica.total.rpf <- Brassica.points[ which(Brassica.points$type == "DNA" &
                                   Brassica.points$treatment == "Rpf+"), ]
Brassica.total.no <- Brassica.points[ which(Brassica.points$type == "DNA" &
                                   Brassica.points$treatment == "Rpf-"), ]

# Add points
# Active community Rpf+
points(Brassica.active.rpf[ ,1], Brassica.active.rpf[ ,2], pch = 21,
       cex = 3.5, col = "Black", bg = "grey15", lwd= 2.5)
# Active community Rpf-
points(Brassica.active.no[ ,1], Brassica.active.no[ ,2], pch = 21,
       cex = 3.5, col = "Black", bg = "lightgrey", lwd= 2.5)   
# Total community Rpf+ 
points(Brassica.total.rpf[ ,1], Brassica.total.rpf[ ,2], pch = 24,
       cex = 3.5, col = "Black", bg = "grey15", lwd= 2.5)
# Total community Rpf-
points(Brassica.total.no[ ,1], Brassica.total.no[ ,2], pch = 24,
       cex = 3.5, col = "Black", bg = "lightgrey", lwd= 2.5)

# Add ellipses 
# Active Rpf+
ordiellipse(cbind(Brassica.active.rpf[ ,1], Brassica.active.rpf[ ,2]), Brassica.active.rpf$treatment, kind="sd", conf=0.95,
            lwd=2, lty=3, draw="lines", col="black", label=FALSE)
# Active Rpf-
ordiellipse(cbind(Brassica.active.no[ ,1], Brassica.active.no[ ,2]), Brassica.active.no$treatment, kind="sd", conf=0.95,
            lwd=2, lty=3, draw="lines", col="black", label=FALSE)
# Total Rpf+
ordiellipse(cbind(Brassica.total.rpf[ ,1], Brassica.total.rpf[ ,2]), Brassica.total.rpf$treatment, kind="sd", conf=0.95,
            lwd=2, lty=3, draw="lines", col="black", label=FALSE)
# Total Rpf- 
ordiellipse(cbind(Brassica.total.no[ ,1], Brassica.total.no[ ,2]), Brassica.total.no$treatment, kind="sd", conf=0.95,
            lwd=2, lty=3, draw="lines", col="black", label=FALSE)

# Add legend outside
par(mar = c(4, 0, 5, 1) + 0.5)
plot.new()
legend(0, 1, c("Active, -Rpf", "Active, +Rpf", "Total, -Rpf", "Total, +Rpf"),
       pch = c(21, 21, 24, 24),
       col = c("Black", "Black", "Black", "Black"),
       pt.bg = c("lightgrey", "grey15", "lightgrey", "grey15"),
       bty = "n", y.intersp = 1, pt.cex = 3.2, cex = 2, lwd= 2, lty =  NA)

# Sample number label
# mtext(text = expression(italic("n")~" = 5"), line = 1, adj = -0.65, col="black", cex = 1.5)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("../figures/FigS5.ActinobacteriaComp.png")
grid.raster(img)
```

# Multivariate test

```{r}
# PERMANOVA - test for main effects and interactions

# Add factor for pot number to account of paired/match design of study
Brassica.div$pot <- c(1,2,3,4,5,6,7,8,9,10,1,2,3,4,5,6,7,8,9,10)
Brassica.div$pot <- as.factor(Brassica.div$pot) # unranked order

# 1) PERMANOVA to test Rpf and metabolic status effects on soil bacterial community structure
all.bray.permanova <- adonis(Actino.OTU.REL.log ~ type + treatment + pot + 
                          type*treatment , data = Brassica.div,
                         method = "bray", binary = FALSE, permutations = 999)
all.bray.permanova # DNA.RNA, P = 0.001; Rpf, P = 0.031; DNA.RNA x Rpf, P = 0.274


# 2) Indicator analysis - identify taxa contributing to differences

# Phi coefficients (actually biserial since we're using abundance data)
# https://cran.r-project.org/web/packages/indicspecies/vignettes/indicspeciesTutorial.pdf

# Treatment code
Rpf.treat <-c(1,1,1,1,1,2,2,2,2,2) # -Rpf then + Rpf

# a) Run phi analysis: DNA
phi.DNA <- multipatt(Actino.OTU.REL.log[11:20,], cluster = Rpf.treat, func = "r.g", 
                     control = how(nperm = 999))

# b) Run phi analysis: RNA
phi.RNA <- multipatt(Actino.OTU.REL.log[1:10,], cluster = Rpf.treat, func = "r.g", 
                     control = how(nperm = 999))

# Subset for significant (p < 0.05) correlations greater than 0.7
phi.sig.DNA <- phi.DNA$sign[ which(phi.DNA$sign[5] <= 0.05 & phi.DNA$sign[4] >= abs(0.7)), ]

# i.) Taxa associated with -RPF treatment 
phi.sig.DNA.minus <- phi.sig.DNA[ which(phi.sig.DNA[1] == 1),] # DNA ~ -Rpf, n = 7
imp.otu.DNA.minus <- as.vector(rownames(phi.sig.DNA.minus))
imp.tax.DNA.minus <- OTU.tax[which (OTU.tax$OTU %in% imp.otu.DNA.minus), ] 
imp.tax.DNA.minus.table <- table(imp.tax.DNA.minus$Genus)
imp.tax.DNA.minus.table.per <- imp.tax.DNA.minus.table/dim(imp.tax.DNA.minus)[1]*100
# Acidothermus, Catenulispora, uncultured, unclassified

# ii.) Taxa associated with +RPF treatment 
phi.sig.DNA.plus <- phi.sig.DNA[ which(phi.sig.DNA[2] == 1),] # DNA ~ +Rpf, n = 0!!!
imp.otu.DNA.plus <- as.vector(rownames(phi.sig.DNA.plus))
imp.tax.DNA.plus <- OTU.tax[which (OTU.tax$OTU %in% imp.otu.DNA.plus), ] 
imp.tax.DNA.plus.table <- table(imp.tax.DNA.plus$Genus)
imp.tax.DNA.plus.table.per <- imp.tax.DNA.plus.table/dim(imp.tax.DNA.plus)[1]*100
# no indicators!

# Subset for significant correlations greater than 0.7
phi.sig.RNA <- phi.RNA$sign[ which(phi.RNA$sign[5] <= 0.05 & phi.RNA$sign[4] >= abs(0.7)), ]

# i.) Taxa associated with -RPF treatment 
phi.sig.RNA.minus <- phi.sig.RNA[ which(phi.sig.RNA[1] == 1),] # RNA ~ -Rpf, n = 3
imp.otu.RNA.minus <- as.vector(rownames(phi.sig.RNA.minus))
imp.tax.RNA.minus <- OTU.tax[which (OTU.tax$OTU %in% imp.otu.RNA.minus), ] 
imp.tax.RNA.minus.table <- table(imp.tax.RNA.minus$Genus)
imp.tax.RNA.minus.table.per <- imp.tax.RNA.minus.table/dim(imp.tax.RNA.minus)[1]*100
# Acidimicrobiaceae_unclassified, Acidothermus, Mycobacterium 

# ii.) Taxa associated with +RPF treatment 
phi.sig.RNA.plus <- phi.sig.RNA[ which(phi.sig.RNA[2] == 1),] # RNA ~ +Rpf, n = 1
imp.otu.RNA.plus <- as.vector(rownames(phi.sig.RNA.plus))
imp.tax.RNA.plus <- OTU.tax[which (OTU.tax$OTU %in% imp.otu.RNA.plus), ] 
imp.tax.RNA.plus.table <- table(imp.tax.RNA.plus$Genus)
imp.tax.RNA.plus.table.per <- imp.tax.RNA.plus.table/dim(imp.tax.RNA.plus)[1]*100
# (Thermoleophilia = Class), Solirubrobacterales unclassified

# iii.) Figure for responders (Solibrobacterales)

# Solibrobacterales taxonomy file
solub.OTU.tax <- OTU.tax[ which(OTU.tax$Genus == 'Solirubrobacterales_unclassified'), ] # n = 704
solub.OTU <- OTU[, solub.OTU.tax$OTU]
# Remove mock community # 
solub.OTU <- solub.OTU[1:20, ]
# Sum counts across OTUs
solub.count <- rowSums(solub.OTU)
actino.count <- rowSums(Actino.OTU)
solub.rel <- solub.count/actino.count


# Summaries for plotting
s.minus.DNA.obs <- solub.rel[11:15]
s.minus.DNA.mean <- mean(solub.rel[11:15])
s.minus.DNA.sem <- sem(solub.rel[11:15])
s.minus.DNA.test <- t.test(s.minus.DNA.obs, conf.level = 0.95)
s.minus.DNA.LL <- s.minus.DNA.test$conf.int[1]
s.minus.DNA.UL <- s.minus.DNA.test$conf.int[2]

s.minus.RNA.obs <- solub.rel[1:5]
s.minus.RNA.mean <- mean(solub.rel[1:5])
s.minus.RNA.sem <- sem(solub.rel[1:5])
s.minus.RNA.test <- t.test(s.minus.RNA.obs, conf.level = 0.95)
s.minus.RNA.LL <- s.minus.RNA.test$conf.int[1]
s.minus.RNA.UL <- s.minus.RNA.test$conf.int[2]
  
s.plus.DNA.obs <- solub.rel[16:20]
s.plus.DNA.mean <- mean(solub.rel[16:20])
s.plus.DNA.sem <- sem(solub.rel[16:20])
s.plus.DNA.test <- t.test(s.plus.DNA.obs, conf.level = 0.95)
s.plus.DNA.LL <- s.plus.DNA.test$conf.int[1]
s.plus.DNA.UL <- s.plus.DNA.test$conf.int[2]

s.plus.RNA.obs <- solub.rel[6:10]
s.plus.RNA.mean <- mean(solub.rel[6:10])
s.plus.RNA.sem <- sem(solub.rel[6:10])
s.plus.RNA.test <- t.test(s.plus.RNA.obs, conf.level = 0.95)
s.plus.RNA.LL <- s.plus.RNA.test$conf.int[1]
s.plus.RNA.UL <- s.plus.RNA.test$conf.int[2]


# Plot plant traits

```{r}
png(filename="../figures/Fig2.PlantTraits.png",
    width = 1200, height = 800, res = 96*2)

par(oma=c(7,3,7,1), mar=c(2,3,3,3.5))

# Panel 1: shoot biomass
solub.fig <- plot(jitter(rep(1, length(s.minus.DNA.obs)), amount = 0.1), 
      s.minus.DNA.obs, ylim = c(0, 0.04), xlim = c(0, 6), pch = 21, 
      col = "lightgrey", bg = "lightgrey", lwd = 3.5, cex = 1.7, yaxt = "n", 
      xaxt = "n", cex.lab = 2, cex.axis = 2, las = 1, ylab = "", xlab = "")
      box(lwd = 2)
      
points(jitter(rep(2, length(s.minus.RNA.obs)), amount = 0.1), s.minus.RNA.obs, 
       pch = 21, bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)

points(jitter(rep(4, length(s.plus.DNA.obs)), amount = 0.1), s.plus.DNA.obs, 
       pch = 21, bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)

points(jitter(rep(5, length(s.plus.RNA.obs)), amount = 0.1), s.plus.RNA.obs, 
       pch = 21, bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)

# Add mean  
points(1, mean(s.minus.DNA.obs), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  
points(2, mean(s.minus.RNA.obs), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)
points(4, mean(s.plus.DNA.obs), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)
points(5, mean(s.plus.RNA.obs), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)

box(lwd = 2)

# Y-axis label # 
mtext(expression('Relative abundance (%'), side = 2,
      outer = FALSE, cex = 1.25, line = 3.5, adj = 0.5)

# Major Axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0", "0.5", "1.0"), at = c(0, 0.5, 1.0))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     at=c(0, 0.5, 1.0), labels = F, tck = -0.05)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("-Rpf", "+Rpf"), at = c(1.5, 4.5))
axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     at = c(1.5, 4.5), labels = F, tck = -0.05)

# Add confidence intervals
arrows(x0 = c(1,2,4,5), y0 = c(s.minus.DNA.mean, s.minus.RNA.mean, 
      s.plus.DNA.mean, s.plus.RNA.mean), y1 = c(s.minus.DNA.LL, s.minus.RNA.LL, 
      s.plus.DNA.LL, s.plus.RNA.LL), angle = 90, length = 0.15, lwd = 2)

arrows(x0 = c(1,2,4,5), y0 = c(s.minus.DNA.mean, s.minus.RNA.mean, 
      s.plus.DNA.mean, s.plus.RNA.mean), y1 = c(s.minus.DNA.UL, s.minus.RNA.UL, 
      s.plus.DNA.UL, s.plus.RNA.UL), angle = 90, length = 0.15, lwd = 2)





arrows(x0 = c(1,2,4,5), y0 = c(s.minus.DNA.mean, s.minus.RNA.mean, 
      s.plus.DNA.mean, s.plus.RNA.mean), y1 = c(s.minus.DNA.mean, s.minus.RNA.mean, 
      s.plus.DNA.mean, s.plus.RNA.mean) + c(s.minus.DNA.sem, s.minus.RNA.sem, 
      s.plus.DNA.sem, s.plus.RNA.sem), angle = 90, length = 0.15, lwd = 2)

# Add SEM (negative)
arrows(x0 = c(1,2,4,5), y0 = c(s.minus.DNA.mean, s.minus.RNA.mean, 
      s.plus.DNA.mean, s.plus.RNA.mean), y1 = c(s.minus.DNA.mean, s.minus.RNA.mean, 
      s.plus.DNA.mean, s.plus.RNA.mean) - c(s.minus.DNA.sem, s.minus.RNA.sem, 
      s.plus.DNA.sem, s.plus.RNA.sem), angle = 90, length = 0.15, lwd = 2)
```
