---
title: "Arabidopsis plate biomass"
author: "Venus Kuo"
date: "March 1, 2018"
output: html_document
---

# 1) Question

To determine whether direct application of Rpf alters biomass. 


# 2) Methods

As a control to confirm the neutral effect of Rpf protein on plant biomass, I plated Arabidopsis thalina seeds on MS agar media for 5 weeks and directly applied Rpf to the plate every week. At the end of the experiment, I took plate pictures and estimated relative plant biomass using Adobe Photoshop 2017 cc. 


# 3) Set up work environment  

```{r, message=FALSE, warning=FALSE}
# Setup work enviroment 
rm(list = ls())
setwd("~/../Github/BrassicaRpf/data")

# Load dependencies # 
package.list <- c('vegan', 'ggplot2', 'plyr') 
for (package in package.list) {
  if (!require(package, character.only=T, quietly=T)) { 
    install.packages(package)
    library(package, character.only=T)
  } }
```


# 4) Describe and load data set

Read text file of relative biomass fo A. thaliana on plates between Rpf+ and Rpf- treatment.

```{r}
# Read text file #
seedling <- read.delim("~/../Github/BrassicaRpf/data/seedlingbiomass.txt", sep = ",", head = TRUE)

# Look at data set # 
#str(seedling)

# Calculate relative biomass # 
seedling$Biomass <- (seedling$BiomassPixel)/(seedling$PlatePixel)
seedling$RelativeBiomass <- (seedling$Biomass)/(seedling$Seedlings)*100
```


# 5) Arabidopsis thaliana seedling biomass

## Bargraph: Seedling biomass with Rpf treatment

```{r}
# Calculate biomass means and standard errors of the means # 
seedling.means.sem <- ddply(seedling, c("Treatment"), summarise,
                   mean=mean(RelativeBiomass), sem=sd(RelativeBiomass)/sqrt(length(RelativeBiomass)))
seedling.means.sem <- transform(seedling.means.sem, lower=mean-sem, upper=mean+sem)

# Bargraph #
seedling.barplot <- ggplot(seedling.means.sem, aes(x=Treatment,y = mean)) + 
  geom_bar(width=0.5, colour="black", fill="white" , position = position_dodge(width = 0.65), stat = "identity", size=1)

seedling.barplot + 
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(0.65), 
                data=seedling.means.sem, width = 0.2, size=1) +
  labs(y="Relative Seedling Biomass") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=18),
        axis.text.x=element_text(colour="black", size =18),
        axis.title=element_text(size=20),
        axis.title.y = element_text(colour="black",  margin = margin(0,15,0,10)),
        axis.title.x=element_blank(),
        panel.border = element_rect(linetype = "solid", colour = "black", size=1.5, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 1.5))

ggsave("../figures/SupplementaryFig1-Arabidopsiscontrol.png", width = 20, height = 15, units = "cm")
```

## Statistics: Seedling biomass with Rpf treatment

```{r}
anova <- aov(seedling$RelativeBiomass ~ seedling$Treatment, data = seedling)
summary(anova) 
TukeyHSD(anova) 
# Results: Rpf did not significantly affect relative biomass (p=0.321, F=1.214, DF=1,5)
```




