---
title: "nutrient"
author: "Venus Kuo"
date: "Feb 26, 2018"
output: html_document
---

# 1) Question

Does Rpf treatment alter soil properties such as pH and nutrients?


# 2) Methods

I measured soil pH, C:N ratio, and phosorous from samples taken from week 5 of the growth chamber experiment.


# 3) Set up work environment  

```{r, message=FALSE, warning=FALSE}
# Setup work enviroment 
rm(list = ls())
setwd("~/../Github/BrassicaRpf/data")

# Load dependencies # 
package.list <- c('vegan', 'ggplot2', 'plyr') 
for (package in package.list) {
  if (!require(package, character.only=T, quietly=T)) { 
    install.packages(package)
    library(package, character.only=T)
  } }
```


# 4) Describe and load data set

- CN: Text file containing the week, treatment, rep number, soil sample amount, and soil N and C amounts.

- SoilpH: Text file containing treatment, rep number, pH

- soilPhos: Text file containing treatment, rep number, inorganic P, organic P

```{r}
# Read C:N ratio text file #
CN <- read.delim("~/../Github/BrassicaRpf/data/CN.txt", sep = ",", head = TRUE)

# Remove blank sample last row # 
CN <- CN[1:15,]

# Week number as factor # 
CN$Week <- as.factor(CN$Week)

# Calculate ratio of CN # 
CN$Ratio <- ((CN$Carbon/CN$Nitrogen)/CN$SampleAmount)

# Relativeize the N and C concetration by sample amount # 
CN$RelN <- (CN$Nitrogen/CN$SampleAmount)*100
CN$RelC <- (CN$Carbon/CN$SampleAmount)*100

ph <- read.delim("~/../Github/BrassicaRpf/data/SoilpH.txt", sep = ",", head = TRUE)
```

# 5) Carbon and nitrogen ratio

## Bargraph: C:N ratios

```{r}
# Subset CN ratio # 
CN <- subset(CN, Week == "6")

# Bargraph of CN ratio # 
ratio.means.sem <- ddply(CN, c("Treatment"), summarise,
                   mean=mean(Ratio), sem=sd(Ratio)/sqrt(length(Ratio)))
ratio.means.sem <- transform(ratio.means.sem, lower=mean-sem, upper=mean+sem)

ratio.barplot<- ggplot(ratio.means.sem, aes(x=Treatment,y = mean)) + 
  geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), stat = "identity", size=1)  

ratio.barplot + 
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(0.65), 
                data=ratio.means.sem, width = 0.2, size=1) +
  labs(y="CN ratio") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=16),
        axis.text.x=element_text(colour="black", size=20),
        axis.title.x=element_blank(),
        axis.title = element_text(size=18),
        axis.title.y = element_text(colour="black",  margin = margin(0,15,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=1.5, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 1.5))

ggsave("../figures/SupplmentaryFigure2-CNratio.png", width = 20, height = 15, units = "cm")
```

## Statistical test: C:N ratio

```{r}
# Two factor ANOVA of ratio CN #
anova <- aov(CN$Ratio ~ CN$Treatment, data = CN)
summary(anova) 
TukeyHSD(anova) 
```


# 6) Soil pH

## Bargraph: Effect of Rpf on soil pH 

```{r}
# Bargraph of CN ratio # 
pH.means.sem <- ddply(ph, c("Treatment"), summarise,
                   mean=mean(pH), sem=sd(pH)/sqrt(length(pH)))
pH.means.sem <- transform(pH.means.sem, lower=mean-sem, upper=mean+sem)

pH.barplot<- ggplot(pH.means.sem, aes(x=Treatment,y = mean)) + 
  geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), stat = "identity", size=1)  

pH.barplot + 
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(0.65), 
                data=pH.means.sem, width = 0.2, size=1) +
  labs(y="Soil pH") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=16),
        axis.text.x=element_text(colour="black", size=20),
        axis.title.x=element_blank(),
        axis.title = element_text(size=18),
        axis.title.y = element_text(colour="black",  margin = margin(0,15,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=1.5, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 1.5))

ggsave("../figures/SupplmentaryFigure3-soilph.png", width = 20, height = 15, units = "cm")
```

## Statistical test: Rpf effect on soil pH

```{r}
# Two factor ANOVA of ratio CN #
anova <- aov(pH ~ Treatment, data = ph)
summary(anova) 
TukeyHSD(anova) 
```




