---
title: "Brassica Plant"
author: "Venus Kuo"
date: "March 1, 2018"
output: pdf_document
---

# 1) Question

How does resuscitating the microbial seed bank affect plant traits?


# 2) Methods

- I performed a factoral experiment testing the effect of Rpf (Rpf-/+) and soil sterilization (live/sterilized) treatment on plant reproductive and   biomass production. 


# 3) Set working environment and load packages

```{r, message=FALSE, warning=FALSE}
# Set working directory #
rm(list = ls())
getwd()
setwd("~/../Github/BrassicaRpf/data") # Specific to Venus's computer

# Require or install packages #
package.list <- c('vegan', 'nlme' ,'data.table', 'plyr', 'reshape', 'grid', 'png', 'car', 'bbmle', 'reshape2', 'ggplot2') 
for (package in package.list) {
  if (!require(package, character.only=T, quietly=T)) { 
    install.packages(package)
    library(package, character.only=T)
  } }

sem <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
```


# 4) Load and describe data set

plantfitness : ID, Treatment (Rpf+/Rpf-), soil type (live/sterilized), flower.count, seed.count, SLA (Specific leaf area), height, abiomass (above ground biomass), bbiomass (below ground biomass), dweight (dry weight), wweight (wet weight), shoot.root (shoot:root ratio), total.biomass

```{r, message=FALSE, warning=FALSE}
setwd("~/../Github/BrassicaRpf/data") 

# Load txt file # 
data <- read.delim("plantfitness.txt", sep = ",", header = TRUE)

# Remove incorrectly handled sample #
data <- data[-33, ] 

# Remove soil sterilization treatment from dataset # 
data <- subset(data, soil == "Live")
```


# 5) Brassica rapa reproductive output figure and statistics

## Bargraph: Seed Production

```{r}
# Calculating standard error #
seed.mean.sem <- ddply(data, c("Treatment"), summarise, 
                   mean=mean(seed.count), sem=sd(seed.count)/sqrt(length(seed.count)))
seed.mean.sem <- transform(seed.mean.sem, lower=mean-sem, upper=mean+sem) 

# Barplot of mean seed counts #
seed.mean.barplot <- ggplot(seed.mean.sem, aes(x=Treatment, y = mean)) + 
  geom_bar(width=0.5, colour = "black", position = position_dodge(width = 0.65), 
           stat = "identity", size=1, fill=c("white", "grey30")) 

seed.mean.barplot + 
  geom_errorbar(aes(ymax=upper, ymin=lower), position = position_dodge(width = 0.65), 
                data=seed.mean.sem, width = 0.2, size = 1) +
  labs(x = "Treatment", y="Seeds Produced/Plant") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=18),
        axis.text.x=element_text(colour="black", size =20),
        axis.title=element_text(size=22),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",  margin = margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2)) #+
#  annotate("text", x =0.575, y=65, label = c("A"), size=12)
```

## Statistics: Seed count

```{r}
# Two-factor ANOVA of the effect of treatment and soil type on seed count # 
seed.aov <- anova(lm(seed.count ~ Treatment, data=data))
seed.aov
# Results: No significant main effect of Rpf or interaction effect p = 0.42, F= 0.69 
```


## Bargraph: Flower Production

```{r}
# Calculating standard error of the mean for flower production # 
flower.mean.sem <- ddply(data, c("Treatment"), summarise,
                   mean=mean(flower.count), sem=sd(flower.count)/sqrt(length(flower.count)))
flower.mean.sem <- transform(flower.mean.sem, lower=mean-sem, upper=mean+sem)

# Barplot of flower production by soil type and treatment
flower.mean.barplot <- ggplot(flower.mean.sem, aes(x=Treatment, y = mean)) + 
  geom_bar(width=0.5, colour = "black", position = position_dodge(width = 0.65), 
           stat = "identity", size=1, fill=c("white", "grey30"))

flower.mean.barplot + 
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(.65), 
                data=flower.mean.sem, width = 0.2, size=1) +
  labs(x = "Treatment" , y="Flowers Produced/Plant") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=18),
        axis.text.x=element_text(colour="black", size =20),
        axis.title=element_text(size=22),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",  margin = margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2)) 
```

## Statistics: Flower count

```{r}
# Two-factor ANOVA of the effect of treatment and soil type on seed count # 
flower.aov <- anova(lm(flower.count ~ Treatment, data=data))
flower.aov
# Results: Marginally nonsignificant effect of Rpf treatment p = 0.0942, F = 3.2234
```

## Make x-y plot with error bars by treatment with individual data points

```{r}
# Flower count data for plotting # 
flower.rpf <- data[ which(data$Treatment == "Rpf+"),]
flower.con <- data[ which(data$Treatment == "Rpf-"),]

# Flower count table #
flower.mean <- aggregate(data$flower.count ~ Treatment, data, mean)
flower.sem <- aggregate(data$flower.count ~ Treatment, data, sem)
flower.sem.LL <- flower.mean[2] + flower.sem[2]   # For SEM 
flower.sem.UL <- flower.mean[2] - flower.sem[2]

#flower.95.LL <- aggregate(flower.count ~ Treatment, data,   # For 95% CI
#          FUN = function(x) t.test(x)$conf.int[1])
#flower.95.UL <- aggregate(flower.count ~ Treatment, data,
#          FUN = function(x) t.test(x)$conf.int[2])
#flower.table <- data.frame(flower.mean[1], flower.mean[2], flower.sem[2],
#          flower.95.LL[2], flower.95.UL[2])

flower.table <- data.frame(flower.mean[1], flower.mean[2], flower.sem[2],
          flower.sem.LL[1], flower.sem.UL[1])
colnames(flower.table) <- c("Treatment", "mean", "sem", "LCI", "UCI")
flower.table <- flower.table[order(flower.table$mean),]

# Generating figure # 
png(filename="../figures/Figure1-Flower.count.png",
    width = 800, height = 800, res = 96*2)

par(mar = c(4, 5, 1, 1))

flowercount.fig <- plot(jitter(rep(1, length(flower.rpf$flower.count)), amount = 0.1), flower.rpf$flower.count, 
      ylim = c(0, 15), xlim = c(0.5, 2.5), pch = 21, col = "lightgrey", bg = "lightgrey", lwd = 3.5,
      cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, cex.axis = 2,
      las = 1, ylab = "", xlab = "")
      box(lwd = 2)
points(jitter(rep(2, length(flower.con$flower.count)), amount = 0.1), flower.con$flower.count, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)

points(1, mean(flower.rpf$flower.count), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(2, mean(flower.con$flower.count), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  

box(lwd = 2)

mtext(expression('Flower Number'), side = 2,
      outer = FALSE, cex = 1.5, line = 3, adj = 0.5)

# Major Axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0", "3", "6", "9", "12", "15"), at = c(0, 3, 6, 9, 12, 15))

axis(side = 4, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     at=c(0, 3, 6, 9, 12, 15), labels = F, tck = -0.02)

axis(side = 1, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     labels = c("Rpf+", "Rpf-"), at = c(1, 2))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     at = c(1, 2), labels = F, tck = -0.02)

arrows(x0 = c(1,2), y0 = flower.table$mean, y1 = flower.table$LCI, angle = 90,
       length = 0.25, lwd = 2)
arrows(x0 = c(1,2), y0 = flower.table$mean, y1 = flower.table$UCI, angle = 90,
       length=0.25, lwd = 2)

text(0.7,14 ,labels="B", col="black", cex=2)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("../figures/Figure1B-Flower.count.png")
grid.raster(img)
```


# 6) Brassica rapa total plant biomass 

## Bargraph: Total Plant Biomass

```{r}
# Calculating standard error of means for total plant biomass #
plantmeans.sem <- ddply(data, c("Treatment"), summarise,
                   mean=mean(total.biomass), sem=sd(total.biomass)/sqrt(length(total.biomass)))
plantmeans.sem <- transform(plantmeans.sem, lower=mean-sem, upper=mean+sem)

# Barplot of total plant biomass #
plantmeans.barplot<- ggplot(plantmeans.sem, aes(x= Treatment , y = mean)) + 
   geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), 
            stat = "identity", size=1, fill=c("white", "grey30"))

plantmeans.barplot + 
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(0.65), 
                data=plantmeans.sem, width = 0.2, size=1) +
  labs(x = "Treatment", y="Total Plant Biomass (g)") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=18),
        axis.text.x=element_text(colour="black", size =20),
        axis.title=element_text(size=22),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,1,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2)) + 
  annotate("text", x =0.575, y=2, label = c("B"), size=12)
```

## Statistics: Total plant biomass

```{r}
# Two-factor ANOVA of total biomass with treatment and soil type # 
plant.t.aov <- anova(lm(total.biomass ~ Treatment, data=data))
plant.t.aov
# Results: Significant main effect of Rpf, p = 0.0132, F=8.0455
```

# 7) Brassica rapa below-ground (root) biomass 

## Bargraph: Root Biomass

```{r}
# Calculating standard error of means for total plant biomass #
plantmeans.sem <- ddply(data, c("Treatment"), summarise,
                   mean=mean(bbiomass), sem=sd(bbiomass)/sqrt(length(bbiomass)))
plantmeans.sem <- transform(plantmeans.sem, lower=mean-sem, upper=mean+sem)

# Plotting bar graph # 
live.biomass.bar <- ggplot(plantmeans.sem, aes(x = Treatment, y = mean)) + 
  geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), 
           stat = "identity", size=1, fill=c("white", "grey30"))

# Plotting bargraph # 
live.biomass.bar  + geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(0.65), 
                                  data=plantmeans.sem, width = 0.2, size=1) +
  labs(x="Treatment", y="Plant Root Biomass (g)") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=18),
        axis.text.x=element_text(colour="black", size =18),
        axis.title=element_text(size=22),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(5,0,0,0)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2)) +
  annotate("text", x =0.575, y=1.055, label = c("C"), size=12)
```

## Stastistics: Root biomass

```{r}
# Two-factor ANOVA of root biomass with treatment and soil type # 
plant.b.aov <- anova(lm(bbiomass ~ Treatment, data=data))
plant.b.aov
# Results: Signficant main effect of Rpf (p=0.037, F=5.298)
```

# 8) Brassica rapa above-ground biomass 

## Bargraph: Above-ground biomass

```{r}
# Calculating standard error of means for total plant biomass #
plantmeans.a.sem <- ddply(data, c("Treatment"), summarise,
                   mean=mean(abiomass), sem=sd(abiomass)/sqrt(length(abiomass)))
plantmeans.a.sem <- transform(plantmeans.a.sem, lower=mean-sem, upper=mean+sem)

# Barplot of total plant biomass #
plantmeans.a.barplot <- ggplot(plantmeans.a.sem, aes(x = Treatment, y = mean)) + 
  geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), stat = "identity", size=1, fill = c("White", "Grey30"))

# Bargraph #
plantmeans.a.barplot + geom_errorbar(aes(ymax=upper, ymin=lower),
                              position=position_dodge(0.65), data=plantmeans.a.sem, width = 0.2, size=1) +
  labs(x="Treatment", y="Above-ground Biomass (g)") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=18),
        axis.text.x=element_text(colour="black", size =18),
        axis.title=element_text(size=22),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(5,0,0,0)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2)) + 
  annotate("text", x =0.575, y=1.025, label = c("D"), size=12)
```


## Stastistics: Above-ground biomass

```{r}
# Two-factor ANOVA of above-ground biomass with treatment and soil type # 
plant.a.aov <- anova(lm(abiomass ~ Treatment, data=data))
plant.a.aov
# Results: Significant effect of treatment p = 0.017, F= 7.3118
```

## Below- and above-ground plant biomass in the same figure

```{r}
# Subsetting data set # 
plant.sub.ab <- subset(data, select=(c(Treatment, abiomass, bbiomass)))

# Combining the above and belowground biomass # 
plant.m <- melt(plant.sub.ab) 

# Summarizing means of plant biomass data # 
plant.m.means <- ddply(plant.m, c("Treatment", "variable"), summarise, mean=mean(value))

# Plotting stacked bar graph # 
live.biomass.bar <- ggplot(plant.m.means, aes(x = variable, y = mean, fill = Treatment)) + 
  geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), stat = "identity", size=1)

# Calculating standard error of means for total plant biomass #
plantmeans.sem <- ddply(plant.m, c("Treatment", "variable"), summarise,
                   mean=mean(value), sem=sd(value)/sqrt(length(value)))
plantmeans.sem <- transform(plantmeans.sem, lower=mean-sem, upper=mean+sem)

# Create label for axis.label.x # 
label <- c("Above-ground", "Below-ground")

# Plotting stacked bargraph # 
live.biomass.bar  + geom_errorbar(aes(ymax=upper, ymin=lower),
                              position=position_dodge(0.65), data=plantmeans.sem, width = 0.2, size=1) +
  labs(x =  "Plant Section" ,y="Plant Biomass (g)") +
  theme_classic() +
  scale_fill_manual(values=c("White", "Grey30")) +
  theme(axis.text.y=element_text(colour="black", size=18),
        axis.text.x=element_text(colour="black", size =20),
        axis.title=element_text(size=22),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2),
        legend.title = element_text(size=18),
        legend.text=element_text(size=18),
        legend.position = c(0.875, 0.85))+
  annotate("text", x =0.575, y=1.05, label = c("B"), size=12) +
  scale_x_discrete(labels = label)
```


## Make x-y plot of biomass change with error bars by treatment with individual data points 

```{r}
# Subsetting data set # 
data.sub <- subset(data, select=(c(Treatment, abiomass, bbiomass)))

# Combining the above and belowground biomass # 
data.m <- melt(data.sub) 

# Biomass count data for plotting # 
root.rpf <- data.m[ which(data.m$Treatment == "Rpf+" & data.m$variable == "bbiomass"),]
root.con <- data.m[ which(data.m$Treatment == "Rpf-" & data.m$variable == "bbiomass"),]
shoot.rpf <- data.m[ which(data.m$Treatment == "Rpf+" & data.m$variable == "abiomass"),]
shoot.con <- data.m[ which(data.m$Treatment == "Rpf-" & data.m$variable == "abiomass"),]

# Plant biomass count table #
root.mean <- aggregate(data$bbiomass ~ Treatment, data, mean)
root.sem <- aggregate(data$bbiomass ~ Treatment, data, sem)
shoot.mean <- aggregate(data$abiomass ~ Treatment, data, mean)
shoot.sem <- aggregate(data$abiomass ~ Treatment, data, sem)
root.sem.LL <- root.mean[2] + root.sem[2]
root.sem.UL <- root.mean[2] - root.sem[2]
shoot.sem.LL <- shoot.mean[2] + shoot.sem[2]
shoot.sem.UL <- shoot.mean[2] - shoot.sem[2]
root.table <- data.frame(root.mean[1], root.mean[2], root.sem[2],
          root.sem.LL[1], root.sem.UL[1])
shoot.table <- data.frame(shoot.mean[1], shoot.mean[2], shoot.sem[2],
          shoot.sem.LL[1], shoot.sem.UL[1])

#root.95.LL <- aggregate(bbiomass ~ Treatment, data,
#          FUN = function(x) t.test(x)$conf.int[1])
#root.95.UL <- aggregate(bbiomass ~ Treatment, data,
#          FUN = function(x) t.test(x)$conf.int[2])
#root.table <- data.frame(root.mean[1], root.mean[2], root.sem[2],
#          root.95.LL[2], root.95.UL[2])
colnames(root.table) <- c("Treatment", "mean", "sem", "LCI", "UCI")
root.table <- root.table[order(root.table$mean),]
#root.table$plant <- c("root", "root")

# Shoot biomass count table # 
#shoot.95.LL <- aggregate(abiomass ~ Treatment, data,
#          FUN = function(x) t.test(x)$conf.int[1])
#shoot.95.UL <- aggregate(abiomass ~ Treatment, data,
#          FUN = function(x) t.test(x)$conf.int[2])
#shoot.table <- data.frame(shoot.mean[1], shoot.mean[2], shoot.sem[2],
#          shoot.95.LL[2], shoot.95.UL[2])
colnames(shoot.table) <- c("Treatment", "mean", "sem", "LCI", "UCI")
shoot.table <- shoot.table[order(shoot.table$mean),]
#shoot.table$plant <- c("shoot", "shoot")

# Combine count table # 
biomass.table <- rbind(root.table, shoot.table)

# Generating figure # 
png(filename="../figures/Figure1A-Plantbiomass.png",
    width = 800, height = 800, res = 96*2)

par(mar = c(5, 5, 1, 1))

biomass.fig <- plot(jitter(rep(1, length(root.rpf$value)), amount = 0.1), root.rpf$value, 
      ylim = c(0, 1.5), xlim = c(0.5, 4.5), pch = 21, col = "lightgrey", bg = "lightgrey", lwd = 3.5,
      cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, cex.axis = 2,
      las = 1, ylab = "", xlab = "")
      box(lwd = 2)
points(jitter(rep(2, length(root.con$value)), amount = 0.1), root.con$value, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)
points(jitter(rep(3, length(shoot.rpf$value)), amount = 0.1), shoot.rpf$value, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)
points(jitter(rep(4, length(shoot.con$value)), amount = 0.1),shoot.con$value, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7) 

points(1, mean(root.rpf$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(2, mean(root.con$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  
points(3, mean(shoot.rpf$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(4, mean(shoot.con$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  

box(lwd = 2)

mtext(expression('Dry Biomass (g)'), side = 2,
      outer = FALSE, cex = 1.5, line = 3, adj = 0.5)
mtext(expression('Root           Shoot'), side = 1, 
      outer = FALSE, cex = 1.5, line = 2.9, adj = 0.5)

# Major Axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("0", "0.3", "0.6", "0.9", "1.2", "1.5"), at = c(0, 0.3, 0.6, 0.9, 1.2, 1.5))
axis(side = 4, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     at=c(0, 0.3, 0.6, 0.9, 1.2, 1.5), labels = F, tck = -0.02)
axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("Rpf+", "Rpf-", "Rpf+", "Rpf-"), at = c(1, 2, 3, 4))
axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     at = c(1, 2, 3, 4), labels = F, tck = -0.02)

arrows(x0 = c(1,2,3,4), y0 = biomass.table$mean, y1 = biomass.table$LCI, angle = 90,
       length = 0.25, lwd = 2)
arrows(x0 = c(1,2,3,4), y0 = biomass.table$mean, y1 = biomass.table$UCI, angle = 90,
       length=0.25, lwd = 2)

text(0.9,1.375 ,labels="A", col="black", cex=2)

abline(v=2.5,col="Black",lty=2, lwd=3)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("../figures/Figure1A-Plantbiomass.png")
grid.raster(img)
```


# 9) Brassica rapa root:shoot ratio

## Bargraph: Root : Shoot Ratio

```{r}
# Root shoot ratio calculations # 
abio <- data$abiomass
bbio <- data$bbiomass
data$root.shoot.ratio <- abio/bbio

# Calculate means and SEM # 
root.shoot.means.sem <- ddply(data, c("Treatment"), summarise,
                   mean=mean(root.shoot.ratio), sem=sd(root.shoot.ratio)/sqrt(length(root.shoot.ratio)))
root.shoot.means.sem <- transform(root.shoot.means.sem, lower=mean-sem, upper=mean+sem)

# Bar plot # 
root.shoot.bar <- ggplot(root.shoot.means.sem, aes(x = Treatment, y = mean)) + 
  geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), 
           stat = "identity", size=1, fill = c("White", "Grey30"))

root.shoot.bar  + geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(0.65), 
                                data=root.shoot.means.sem, width = 0.2, size=1) +
  labs(x =  "Treatment" ,y="Root:Shoot Ratio") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=18),
        axis.text.x=element_text(colour="black", size =20),
        axis.title=element_text(size=22),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2))# + 
 # annotate("text", x =0.575, y=2.3, label = c("D"), size=12)
```

## Statistics: Root : Shoot Ratio

```{r}
ratiobiomass.aov <- anova(lm(root.shoot.ratio~Treatment ,data=data))
ratiobiomass.aov
# Result: Non-significant main effect of Rpf 
```


# 10) Plant specific leaf area 

## Bargraph: SLA

```{r}
# Calculate means and SEM # 
SLA.sem <- ddply(data, c("Treatment"), summarise,
                   mean=mean(SLA), sem=sd(SLA)/sqrt(length(SLA)))
SLA.sem <- transform(SLA.sem, lower=mean-sem, upper=mean+sem)

# Bar plot # 
SLA.bar <- ggplot(SLA.sem, aes(x = Treatment, y = mean)) + 
  geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), stat = "identity", size=1, fill = c("white", "grey30"))

SLA.bar  + geom_errorbar(aes(ymax=upper, ymin=lower),
                              position=position_dodge(0.65), data=SLA.sem, width = 0.2, size=1) +
  labs(x =  "Treatment" ,y="Plant SLA") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=18),
        axis.text.x=element_text(colour="black", size =20),
        axis.title=element_text(size=22),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2)) +
  annotate("text", x =0.575, y=300, label = c("D"), size=12)
```

## Statistics: Plant SLA

```{r}
SLA.aov <- anova(lm(SLA~Treatment, data=data))
SLA.aov
# Results: Non-significant effec of Rpf p = 0.4355, f=0.6446
```

# 11) Plant Height

## Bargraph: SLA

```{r}
# Calculate means and SEM # 
Height.sem <- ddply(data, c("Treatment"), summarise,
                   mean=mean(height), sem=sd(height)/sqrt(length(height)))
Height.sem <- transform(Height.sem, lower=mean-sem, upper=mean+sem)

# Bar plot # 
Height.bar <- ggplot(Height.sem, aes(x = Treatment, y = mean)) + 
  geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), stat = "identity", size=1, fill=c("White", "grey30"))

Height.bar  + geom_errorbar(aes(ymax=upper, ymin=lower),
                              position=position_dodge(0.65), data=Height.sem, width = 0.2, size=1) +
  labs(x =  "Treatment" ,y="Plant Shoot Height") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=18),
        axis.text.x=element_text(colour="black", size =20),
        axis.title=element_text(size=22),
        axis.title.y = element_text(colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(colour="black",margin=margin(15,10,0,10)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2)) + 
  annotate("text", x =0.575, y=30, label = c("D"), size=12)
```

## Statistics: Plant shoot height

```{r}
height.aov <- anova(lm(height~Treatment ,data=data))
height.aov
# Results: Non-significant Rpf effect p=0.2856, F=1.2325
```










