---
title: "Soil Abundance Activity"
author: "Venus Kuo"
date: "March 1, 2018"
output: html_document
---
# 1) Question

How does resuscitating the microbial seed bank affect soil microbial community activity and abundance?


# 2) Methods

I extracted soil samples throughout the six-week growth chamber experiment to measure soil respiration and soil microbial (bacteria/fungal) abundance 


# 3) Set working environment and load packages

```{r, message=FALSE, warning=FALSE}
# Set working directory #
rm(list = ls())
getwd()
setwd("~/../Github/BrassicaRpf/data")

# Require or install packages #
package.list <- c('plyr', 'grid' ,'png', 'car', 'bbmle', 'ggplot2', 'vegan', 'tidyr', 'dplyr', 'codyn', 'cowplot', 'png', 'MullerPlot', 'RColorBrewer', 'pander', 'reshape', 'nlme') 
for (package in package.list) {
  if (!require(package, character.only=T, quietly=T)) { 
    install.packages(package)
    library(package, character.only=T)
  } }

# Load small function # 
sem <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}

# Fancy Scientific function # 
fancy_scientific <- function(l) {
     # turn in to character string in scientific notation
     l <- format(l, scientific = TRUE)
     # quote the part before the exponent to keep all the digits
     l <- gsub("^(.*)e", "'\\1'e", l)
     # turn the 'e+' into plotmath format
     l <- gsub("e", "%*%10^", l)
     # return this as an expression
     parse(text=l)
}
```


# 4) Load data 

qPCR.txt : Bacterial (EUB 338f-515r) and fungal (5.8f-ITs4r) gene copy abundances in live soil DNA taken from experimental pots containing Brassica plants. Treatment, Rep, Week, Well position, and Gene copy amount.

GCH_CO2.txt : Soil respiration dataset. ID, Treatment (Rpf+/Rpf-), Soil (Live/Sterilized), plant (present/absent), hour, week 1-6.

```{r , message=FALSE, warning=FALSE}
# Set working directory # 
setwd("~/../Github/BrassicaRpf/data")

# Load microbial qPCR abundance data # 
abundance <- read.csv("qPCR.txt", sep = ",", header = TRUE) 

# Load soil respirtation data #
CO2 <- read.csv("GCH_CO2.txt", sep = ",", header = TRUE)

# Define week as factor for qPCR data # 
abundance$Week<- factor(abundance$Week)

# Remove samples with NA # 
abundance <- na.omit(abundance)

# Standardizing gene copy abundances by soil amount # 
soil <- abundance$Soil
abundance$stdGeneCopy <- (abundance$GeneCopy)/soil

# Log10 transforming standardized gene copy number # 
abundance$log10stdGeneCopy <- log10(abundance$stdGeneCopy)

# Subset week 6 # 
abundance <- subset(abundance, Week == "6")
```


# 5) Soil microbial activity

Soil CO2 respiration by Rpf and sterilization treatment

## Line graph: Soil respiration

```{r}
# Subsetting data from CO2 dataset # 
CO2.sub <- subset(CO2, plant=="present" & hour =="24" & soil=="live", select=c(Treatment, soil, Week.1, Week.2, Week.3, Week.4, Week.5, Week.6))

# Change column names # 
colnames(CO2.sub) <- c("Treatment", "soil", "1","2","3","4","5","6")

# Melt data into three columns of treatment, soil and weeks # 
CO2.m <- melt(CO2.sub)
CO2.m$variable <- as.factor(CO2.m$variable)

# Standardizing the y-axis units by 24 hours # 
CO2.values <- CO2.m$value
CO2.m$StdCO2 <- CO2.values/24

# Calculating mean and sem # 
CO2.means.sem <- ddply(CO2.m, c("Treatment", "variable"), summarise,
                   mean=mean(StdCO2), sem=sd(StdCO2)/sqrt(length(StdCO2)))
CO2.means.sem <- transform(CO2.means.sem, lower=mean-sem, upper=mean+sem)

# The errorbars overlapped, so use position_dodge to move them horizontally
pd <- position_dodge(0)

# Plot line graph # 
ggplot(CO2.means.sem, aes(x=variable, y=mean, colour=Treatment, group=Treatment)) + 
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(0), 
                data=CO2.means.sem, width = 0.2, size=1) +
    geom_line(position=pd, size=2) +
    geom_point(aes(shape=Treatment), position=pd, size=6, fill="White") +
    scale_shape_manual(values=c(16, 17)) +
    xlab("Time (weeks)") +
    ylab(expression(Soil~Respiration~(ppm~CO[2]~d^-1))) +
    theme_classic() +
    theme(axis.text.y=element_text(colour="black", size=18),
          axis.text.x=element_text(colour="black", size=20),
          axis.title.y = element_text(size = 19, colour="black",  margin = margin(0,10,0,0)),
          axis.title.x = element_text(size = 20,colour="black",margin=margin(15,10,0,10)),
          panel.border = element_rect(linetype = "solid", colour = "black", size=2, fill=NA),
          legend.position = c(0.9, 0.2), legend.text = element_text(size=15),
          legend.title = element_text(size=15)) +
    scale_color_manual(values=c('Black','Black')) 

ggsave("../figures/Figure2-Soilrespiration.png", width = 20, height = 15, units = "cm")
```

## Statistics: Soil microbial respiration 

```{r}
# Perform RM-ANOVA #   
CO2.rm <- lme(value ~ variable*Treatment, random = ~ 1 | soil,
              correlation = corAR1(form = ~1 | soil), 
              data = CO2.m)

rich.cmp <- lme(value ~ variable*Treatment, random = ~ 1 | soil,
                correlation = corCompSymm(form = ~1 | soil),
                data = CO2.m)

#summary(rich.cmp)
#summary(CO2.rm)

# Make cleaner ANOVA table #
set.caption("RMANOVA for soil CO2 respiration")
pander(anova(rich.cmp))

# Make cleaner ANOVA table #
set.caption("RMANOVA for soil CO2 respiration")
pander(anova(CO2.rm))
```


# 6) Soil microbial abundance

## Bargraph: Bacterial 16S rRNA gene copy abundance by Rpf 

```{r}
# Subset bacterial 16S dataset # 
bac.abundance <- subset(abundance, Gene=="rRNA" & Week == "6", select=c(Treatment, Week, stdGeneCopy))

# Calculate bacterial gene abundance standard errors of the means # 
bac.means.sem <- ddply(bac.abundance, c("Treatment"), summarise,
                   mean=mean(stdGeneCopy), sem=sd(stdGeneCopy)/sqrt(length(stdGeneCopy)))
bac.means.sem <- transform(bac.means.sem, lower=mean-sem, upper=mean+sem)

# Barplot of bacterial gene abundance in soil # 
bac.means.barplot <- ggplot(bac.means.sem, aes(x=Treatment, y = mean)) + 
  geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), stat = "identity", size=1) 

bac.means.barplot +
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(0.65), 
                data=bac.means.sem, width = 0.2, size=1) +
  labs(x="Treatment", y="16S rRNA Gene Copy/Soil (g)") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=16),
        axis.text.x=element_text(colour="black", size =20),
        axis.title.y = element_text(size = 19, colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(size = 20, colour="black",margin=margin(15,0,0,0)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2)) +
  scale_y_continuous(labels=fancy_scientific)
```

## Statistics: Bacterial abundance by Rpf treatment and week

```{r}
# Bacterial 16S rRNA # 
bac.aov <- aov(stdGeneCopy ~ Treatment, data=bac.abundance)
summary(bac.aov)
TukeyHSD(bac.aov)
# Results: Significant effect of Rpf p=0.01, F= 7.407, DF=1,34 
```


## Bargraph: Fungal ITS gene abundance by Rpf and week

```{r}
# Fungal ITS # 
# Subset ITS gene copy data # 
ITS.abundance <- subset(abundance, Gene=="ITS" & Week == "6" , select=c(Treatment, stdGeneCopy))

# Calculate fungal ITS gene abundance standard errors of the means # 
ITS.means.sem <- ddply(ITS.abundance, c("Treatment"), summarise,
                   mean=mean(stdGeneCopy), sem=sd(stdGeneCopy)/sqrt(length(stdGeneCopy)))
ITS.means.sem <- transform(ITS.means.sem, lower=mean-sem, upper=mean+sem)

# Barplot of Fungal ITS abundance in soil # 
ITS.mean.barplot<- ggplot(ITS.means.sem, aes(x=Treatment ,y = mean)) +
  geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), stat = "identity", size=1) 

# Barplot of fungal gene abundance in soil for pub figures # 
ITS.mean.barplot + 
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(0.65), 
                data=ITS.means.sem, width = 0.2, size=1) +
  labs(x="Treatment", y="ITS Gene Copy/Soil (g)") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=16),
        axis.text.x=element_text(colour="black", size =20),
        axis.title.y = element_text(size = 19, colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(size = 20, colour="black",margin=margin(15,0,0,0)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2)) +
  scale_y_continuous(labels=fancy_scientific) 
```

## Statistics: Fungal abundance by Rpf treatment and week

```{r}
# Fungal ITS gene # 
# Two Factor ANOVA of ITS Fungal gene copy abundances affected by Treatment and week of growth chamber experiment # 
fun.aov <- aov(stdGeneCopy ~ Treatment, data=ITS.abundance)
summary(fun.aov)
TukeyHSD(fun.aov)
# Results: Significant effect of Rpf on fungi abundance (p=0.0048,F=10.71,df=1,16) # 
```

## Make x-y plot of microbial abundance with error bars by treatment with individual data points 

```{r}
# Subsetting data set # 
data.sub <- subset(abundance, select=(c(Treatment, Gene, stdGeneCopy)))

data.sub.bac <- subset(data.sub, Gene == "rRNA")
data.sub.fun <- subset(data.sub, Gene == "ITS")

# Combining the above and belowground biomass # 
data.m <- melt(data.sub) 

# Biomass count data for plotting # 
bac.rpf <- data.m[ which(data.m$Treatment == "Rpf+" & data.m$Gene == "rRNA"),]
bac.con <- data.m[ which(data.m$Treatment == "Rpf-" & data.m$Gene == "rRNA"),]
fun.rpf <- data.m[ which(data.m$Treatment == "Rpf+" & data.m$Gene == "ITS"),]
fun.con <- data.m[ which(data.m$Treatment == "Rpf-" & data.m$Gene == "ITS"),]

# Abundance count table #
bac.mean <- aggregate(data.sub.bac$stdGeneCopy ~ Treatment, data.sub.bac, mean)
bac.sem <- aggregate(data.sub.bac$stdGeneCopy ~ Treatment, data.sub.bac, sem)

fun.mean <- aggregate(data.sub.fun$stdGeneCopy ~ Treatment, data.sub.fun, mean)
fun.sem <- aggregate(data.sub.fun$stdGeneCopy ~ Treatment, data.sub.fun, sem)

#bac.sem.LL <- bac.mean[2] + bac.sem[2] # For sem
#bac.sem.UL <- bac.mean[2] - bac.sem[2]
#fun.sem.LL <- fun.mean[2] + fun.sem[2]
#fun.sem.UL <- fun.mean[2] - fun.sem[2]
#bac.table <- data.frame(bac.mean[1], bac.mean[2], bac.sem[2],
#          bac.sem.LL[1], bac.sem.UL[1])
#fun.table <- data.frame(fun.mean[1], fun.mean[2], fun.sem[2],
#          fun.sem.LL[1], fun.sem.UL[1])

bac.95.LL <- aggregate(stdGeneCopy ~ Treatment, data.sub.bac, # For 95% CI 
          FUN = function(x) t.test(x)$conf.int[1])
bac.95.UL <- aggregate(stdGeneCopy ~ Treatment, data.sub.bac,
          FUN = function(x) t.test(x)$conf.int[2])
bac.table <- data.frame(bac.mean[1], bac.mean[2], bac.sem[2],
          bac.95.LL[2], bac.95.UL[2])
# fun abundance count table # 
fun.95.LL <- aggregate(stdGeneCopy ~ Treatment, data.sub.fun,
          FUN = function(x) t.test(x)$conf.int[1])
fun.95.UL <- aggregate(stdGeneCopy ~ Treatment, data.sub.fun,
          FUN = function(x) t.test(x)$conf.int[2])
fun.table <- data.frame(fun.mean[1], fun.mean[2], fun.sem[2],
          fun.95.LL[2], fun.95.UL[2])

colnames(bac.table) <- c("Treatment", "mean", "sem", "LCI", "UCI")
bac.table <- bac.table[order(bac.table$mean),]

colnames(fun.table) <- c("Treatment", "mean", "sem", "LCI", "UCI")
fun.table <- fun.table[order(fun.table$mean),]

# Combine count table # 
abundance.table <- rbind(bac.table, fun.table)
abundance.table <- abundance.table[c(1,2,4,3),] 

# Generating figure # 
png(filename="../figures/Figure3-Microbialabundance.CI.png",
    width = 800, height = 800, res = 96*2)

#par(mar = c(5, 5, 1, 1))

par(mar=c(5,5,1,1))

biomass.fig <- plot(jitter(rep(1, length(bac.rpf$value)), amount = 0.1), bac.rpf$value, 
      ylim = c(0, 2.5E6), xlim = c(0.5, 4.5), pch = 21, col = "lightgrey", bg = "lightgrey", lwd = 3.5,
      cex = 1.7, yaxt = "n", xaxt = "n", cex.lab = 2, cex.axis = 2,
      las = 1, ylab = "", xlab = "")
      box(lwd = 2)
points(jitter(rep(2, length(bac.con$value)), amount = 0.1), bac.con$value, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)
points(jitter(rep(3, length(fun.rpf$value)), amount = 0.1), fun.rpf$value, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7)
points(jitter(rep(4, length(fun.con$value)), amount = 0.1),fun.con$value, pch = 21, 
       bg = "lightgrey", col = "lightgrey", lwd = 2, cex = 1.7) 

points(1, mean(bac.rpf$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(2, mean(bac.con$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  
points(3, mean(fun.rpf$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5) 
points(4, mean(fun.con$value), pch = 21, col = "black", 
       bg = "NA", lwd = 2, cex = 2.5)  

box(lwd = 2)

mtext(expression('Gene Copy Number/ Soil (g)'), side = 2,
      outer = FALSE, cex = 1.5, line = 3.5, adj = 0.5)
mtext(expression('Bacteria          Fungi'), side = 1, 
      outer = FALSE, cex = 1.5, line = 3, adj = 0.5)

# Major Axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1, 
     labels = c("0.0E0", "5.0E5", "1.0E6", "1.5E6", "2.0E6", "2.5E6"), 
     at = c(0.0E0, 5.0E5, 1.0E6, 1.5E6, 2.0E6, 2.5E6))

axis(side = 4, lwd.ticks = 2, cex.axis = 1, las = 1,
     at=c(0.0E0, 5.0E5, 1.0E6, 1.5E6, 2.0E6, 2.5E6), labels = F, tck = -0.02)

axis(side = 1, lwd.ticks = 2, cex.axis = 1.25, las = 1,
     labels = c("Rpf+", "Rpf-", "Rpf+", "Rpf-"), at = c(1, 2, 3, 4))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
     at = c(1, 2, 3, 4), labels = F, tck = -0.02)

arrows(x0 = c(1,2,3,4), y0 = abundance.table$mean, y1 = abundance.table$LCI, angle = 90,
       length = 0.25, lwd = 2)
arrows(x0 = c(1,2,3,4), y0 = abundance.table$mean, y1 = abundance.table$UCI, angle = 90,
       length=0.25, lwd = 2)

abline(v=2.5,col="Black",lty=2, lwd=3)

# Close Plot Device
dev.off()
graphics.off()

# Show Plot
img <- readPNG("../figures/Figure3-Microbialabundance.CI.png")
grid.raster(img)
```


# 7) Fungi: Bacteria ratios 

## Calculate F:B ratio 

```{r}
# Calculating F : B ratios #
gca.B <- bac.abundance$stdGeneCopy
gca.F <- ITS.abundance$stdGeneCopy
abundance$FB <- (gca.F/gca.B)
```

## Plot F:B ratio

```{r}
# Calculating mean and sem # 
FB.means.sem <- ddply(abundance, c("Treatment"), summarise,
                   mean=mean(FB), sem=sd(FB)/sqrt(length(FB)))
FB.means.sem <- transform(FB.means.sem, lower=mean-sem, upper=mean+sem)

# Barplot of Fungal FB abundance in soil # 
FB.mean.barplot<- ggplot(FB.means.sem, aes(x=Treatment ,y = mean)) +
  geom_bar(width=0.5, colour="black", position = position_dodge(width = 0.65), stat = "identity", size=1) 

# Barplot of fungal gene abundance in soil for pub figures # 
FB.mean.barplot + 
  geom_errorbar(aes(ymax=upper, ymin=lower), position=position_dodge(0.65), 
                data=FB.means.sem, width = 0.2, size=1) +
  labs(x="Treatment", y="F : B ratio") +
  theme_classic() +
  theme(axis.text.y=element_text(colour="black", size=16),
        axis.text.x=element_text(colour="black", size =20),
        axis.title.y = element_text(size = 19, colour="black",  margin = margin(0,10,0,10)),
        axis.title.x = element_text(size = 20, colour="black",margin=margin(15,0,0,0)),
        panel.border = element_rect(linetype = "solid", colour = "black", size=3, fill=NA),
        axis.ticks.length = unit(.25, "cm"),
        axis.ticks = element_line(size = 2)) 

ggsave("../figures/Figure4-FBratio.png", width = 20, height = 15, units = "cm")

```

## Statistic: F:B ratio

```{r}
# F:B ratio #
# One factor ANOVA to test effect of Rpf on F:B ratio # 
FB.aov <- aov(FB ~ Treatment, data=abundance)
summary(FB.aov)
TukeyHSD(FB.aov)
# Results: Significant effect of Rpf on F:B ratio (F=17.2, p =0.0002, df = 1,34) # 
x <- (0.61-0.16)/0.16
x
```










